\newcommand\N{\ensuremath{\mathbb{N}}}
\newcommand\R{\ensuremath{\mathbb{R}}}
\newcommand\Z{\ensuremath{\mathbb{Z}}}
\renewcommand\O{\ensuremath{\emptyset}}
\newcommand\Q{\ensuremath{\mathbb{Q}}}
\newcommand\C{\ensuremath{\mathbb{C}}}
\let\implies\Rightarrow
\let\impliedby\Leftarrow
\let\iff\Leftrightarrow
\let\epsilon\varepsilon
\documentclass[12pt]{exam}
\usepackage{amsmath,amstext,amssymb,amsthm}   
\usepackage{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 

\begin{document}

 

 \begin{center}
 {\Large Homework 1, Math 4150 }
 \\
 \end{center}


\begin{questions}
\question  Exercise Set 1.1, \#6. 
\begin{parts}
\part Let $a,b, c\in \mathbb{Z} $ with $c\neq 0$. Show that  $a\mid b $ if and only if $a c \mid bc $. 

\part Provide a counterexample showing that the conclusion above does not hold if $c =0$. 
\end{parts}

\textbf{Solution.}

(a).
    First we show that $a | b$ implies  $ac | bc$. If $a | b$ then  there exists some $x \in \Z$ such that  $ax = b$. Now multiply $c$ on both sides to get $acx = bc$. This means that there is some $x \in \Z$ for which  $ac$ multiplied by  $x$ is $bc$. In other words by definition we have  $ac | bc$.


    Now we show that $ac | bc$ implies that  $a | b$. If  $ac | bc$, by definition we have  some $x \in \Z$ such that  $ac x = bc$. Now, because  $c \ne 0$ we can divide  $c$ from both sides to get  $ ax = b$. Again by definition this means that  $b$ is a multiple of  $a$ or that $a | b$.


(b). Consider $a = 3$ and  $b = 5$. If  $c = 0$ we have  $ac = 0 $ and  $bc = 0$. Now  $0 | 0$ is true because for any choice of  $x \in \Z$ we have  $0x = 0$. So we have  $ac | bc$. However, we can easily see that  $3 | 5$ is not true. So this counterexample shows that the statement if and only if doesn't hold if  $c = 0$





\newpage 
\question  Exercise Set 1.1, \#9.  
Let $a, m,n$ be positive integers, with $a>1$. Show that $a ^{m}-1 \mid a ^{n}-1$ if and only if $m\mid n$. 

\textbf{Solution.}

First we show that if $m | n$ then  $a^{m} - 1 | a^{n} - 1$. If $m | n$ then we have for some  $x \in \Z$ that  $mx = n$ which means that  $a^{mx} - 1 = a^{n} - 1$. We can write the left hand side as $(a^{m})^{x} - 1 = (a^{m} - 1)(a^{m}^{(x - 1)}+a^{m}^{(x - 2)}+ \dots+ 1) = (a^{m} - 1)(k)$ where $k = (a^{m}^{(x - 1)}+a^{m}^{(x - 2)}+ \dots+ 1)$. This give us, 
$$ (a^{m} - 1) k = a^{n} - 1 $$ 

Or that $a^{m} - 1 | a^{n} - 1$

Now, we show that  $a^{m} - 1 | a^{n} - 1 $ implies that $m | n$. If $a^{m} - 1 | a^{n} - 1$ then that means $\exists x\in \Z$ such that  $(a^{m} - 1)x = a^{n} - 1$. Now as $n > m$ (we know this because $x$ is a positive integer, which means that $a^{n} > a^{m}$ which means that $n > m$) we can take $n = qm + r$ for some  $q,r \in \N$ where $r < m$. Now we can write, 
\begin{align*}
    a^{n} - 1 &= a^{qm}a^{r}- 1\\
              &= a^{qm}a^{r} - a^{r} + a^{r} - 1\\
              &= (a^{qm} - 1)a^{r}+ (a^{r} - 1)\\
\end{align*}

Now we know that $a^{m} - 1 | a^{n} - 1$ by assupmtion and we also know from the above proof that $a^{m} - 1 | a^{qm} - 1$ as $m | qm$. Hence, this also must mean that  $a^{m} - 1  | a^{r} - 1$. However, by construction we have $r < m$ based on how we constructed  $n$ as $n = qm + r$. This meanss that $a^{m} - 1 > a^{r} - 1$. A larger mumber divides a smaller number only when the smaller number is zero. SO we have $a^{r} - 1 = 0$. Or that $r = 0$. This gives us,  $n = qm + r = qm + 0 = qm$ which implies that  $m | n$.
\



\newpage 
\question  Exercise Set 1.1, \#10. Let $n\in \mathbb{Z} 	$.
\begin{parts}
\part  Prove that $3\mid n ^{3}-n$ 
\part  Prove that $5\mid n ^{5}-n$ 
\part  Is it true that $4 \mid n ^{4}-n$? Provide a proof or a counterexample. 
\end{parts}

\textbf{Solution.} 

(a). First we rewrite $n^{3} - n = n(n^2 - 1) = n(n + 1)(n - 1)$

We have three cases, 

\textbf{Case 1:} $n = 3q + 0$ for some  $q \in Z$

 Here $n  = 3q$ so we have  $$n(n + 1)(n - 1) = 3q(3q + 1)(3q - 1) = 3(k)$$  where $k = q(3q + 1)(3q - 1) $. Hence we have $3 | n^{3} - n$

\textbf{Case 2:} $n = 3q + 1$ for some  $q \in Z$

 Here $n = 3q + 1$ so we have  $$n^{3} -n = n(n + 1)(n - 1) = (3q + 1 )(3q + 2)(3q) = 3k$$ where $k = (3q + 1 )(3q + 2)q$ so we have $3 | n^{3} - n$

\textbf{Case 3:} $n = 3q + 2$ for some  $q \in Z$

 Here $n = 3q + 2$ so we have  $$n^{3} - n = n (n + 1)(n - 1) = (3q + 2)(3q + 3)(3q + 1) = 3 (3q + 2)(q + 1)(3q + 1) = 3k$$ where $k =(3q + 2)(q + 1)(3q + 1)  $ which means that $3 | n^{3} - n$

So, in all three cases we have $3 | n^{3} - n$

(b). Similar to above we have 5 cases as any number can only leave reminaders $0, 1, 2, 3, 4$ when divided by 5. We can also expand $$n^{5} - n = n(n^{4} - 1) = n(n^2 + 1)(n + 1)(n - 1)$$


Now the five cases are,


\textbf{Case 1:}  $n = 5q + 0$

 Here $n = 5q$ so  $$n^{5} - n = n(n^2 + 1)(n + 1)(n - 1) = 5q(5q^2 + 1)(5q + 1)(5q - 1) = 5k$$ where $k = q(5q^2 + 1)(5q + 1)(5q - 1)$ which gives us $5 | n^{5} - n$

\textbf{Case 2:}  $n = 5q + 1$

 Here $n = 5q + 1$ so  $$n^{5} - n = n(n^2 + 1)(n + 1)(n - 1) = (5q + 1)((5q + 1)^2 + 1)(5q + 2)(5q) = 5k$$ where $k =  (5q + 1)((5q + 1)^2 + 1)(5q + 2)q$ which gives us $5 | n^{5} - n$

\textbf{Case 3:}  $n = 5q + 2$

 Here $n = 5q + 2$ so  
 \begin{align*}
     n^{5} - n = n(n^2 + 1)(n + 1)(n - 1) &= (5q + 2)((5q + 2)^2 + 1)(5q  + 3)(5q + 1)\\
                                          &= (5q + 2)(25q^2 + 4 + 20q + 1)(5q + 3)(5q + 1)\\
                                          &= (5q + 2)5(5q^2 + 1 + 4q)(5q + 3)(5q + 1)\\
                                          &= 5k
 \end{align*}
 where $k = (5q + 2)(5q^2 + 1 + 4q)(5q + 3)(5q + 1)$ so we have  $5 | n^{5} - n$

 \textbf{Case 4:}  $n = 5q + 3$

 Here $n = 5q + 3$ so  
 \begin{align*}
     n^{5} - n = n(n^2 + 1)(n + 1)(n - 1) &= (5q + 3)((5q + 3)^2 + 1)(5q + 4)(5q + 2) \\
                                          &=  (5q + 3)(25q^2 + 9 + 30q + 1)(5q + 4)(5q + 2) \\
                                          &=  (5q + 3)(25q^2 + 10 + 30q)(5q + 4)(5q + 2) \\
                                          &=  5(5q + 3)(5^2 + 2 + 6)(5q + 4)(5q + 2) \\
                                          &=  5k
 \end{align*}

 where $k= (5q + 3)(5^2 + 2 + 6)(5q + 4)(5q + 2) $ which gives us  $5 | n^{5} - n$


 \textbf{Case 5:}  $n = 5q + 4$

 Here $n = 5q + 4$ so  
 \begin{align*}
     n^{5} - n = n(n^2 + 1)(n + 1)(n - 1) &= (5q + 4)((5q + 4)^2 + 1)(5q + 5)(5q + 3)\\
                                          &=  5(5q + 4)((5q + 4)^2 + 1)(q + 1)(5q + 3)\\
                                          &= 5k
 \end{align*}
 where $k = (5q + 4)((5q + 4)^2 + 1)(q + 1)(5q + 3)$ which means that $5 | n^{5} - n$


 So in all three cases we have $5 | n^{5} - n$





In all cases we have $5 | n^{5} - n$

(c). We need to either prove or disprove that $4 | n^{4} - n$. We will disprove the statement by giving a counter example. Consider $n = 3$. Here we have  $n^{4} - n = 81 - 3 = 78$. We see that $78 = 4 \times 19 + 1$ which means that $4 \not | 78 $ and hence disproves the statement.




\newpage 
\question  Exercise Set 1.2, \#28  \\ 
Let $a,n$ be positive integers, with $a>1$. Show that if $a ^{n}+1$ is prime, then $a $ is even, and $n$ is a power of $2$. 

\textbf{Solution.} 

Let us assume the contrary that $a$ is odd or $n$ is not a power of 2. 

Case 1 we have $n$ is odd. This means that $a^{n}$ is odd (as odd times odd is always odd). If $a^{n}$ is odd then $a^{n} + 1$ is even and not equal to 2(as $a > 1$ we have $a^{n} + 1 > 2$). And we know that 2 is the only even prime number. This means that $a^{n} + 1$ is composite which breaks our assumption that it is prime. Hence, $n$ cannot be odd.

Case 2 we have $n$ not a power of 2. This means that $n$ has an odd factor. So let $r$ be an odd number such that $n = r s$ for some $s \in \Z$. So we have $a^{rs} + 1 = (a^{s})^{r} + 1 = x^{r} + 1^{r}$.  Now as  $r$ is odd we can factorize this as follows, 
$$ x^{r} + 1^{r} =  (x + 1)(x^{r - 1} - x^{r - 2} + x^{r - 3} \dots )$$ 

Now that $x + 1 > 1$ we have $(x + 1) | x^{r} + 1$ implying that $x^{r} + 1$ is not a prime number as it has a prime divisor.  Hence, contradicting our assumption 

So we show that either $a$ has to be even or  $n$ is a power of 2.


\newpage 
\question Let $n >1$ be an  integer. Prove that, if
$ n ^2   + 1 $ is a prime number, then $n^2 + 1$ is expressible in the form $4k + 1$ with integer $k$. 

\textbf{Solution. } 

    Given that $n^2 + 1$ is prime. We know from above that $n$ has to be even as if its odd the number would be composite. Now if $n$ is even then there is some $m \in \Z$ such that  $n = 2m$. Plugging this into  $n^2 + 1$ we have $(2m)^2 + 1 = 4m^2 + 1 = 4k + 1$ if we take $m^2 = k$. Hence, we show that any prime number of the form $n^2 + 1$  can be written in the form $4k + 1$.




\newpage 
\question  Exercise Set 1.3, \#43.  
\begin{parts}
\part Let $a,b,c \in \mathbb{Z} $ with $(a,b)=1$.  If $a \mid c$ and $b\mid c$ prove that $ab\mid  c$. 
\part Provide a counterexample to show why the statement in (a) cannot hold if $(a,b)>1$. 
\part Let $a_1, \ldots, a_n \in \mathbb{Z} $ be pairwise relatively prime numbers. Prove that if each $a_j \mid c$, 
then $a_1 \cdots a_n \mid c$. 
\end{parts}

\textbf{Solution. } 

(a). We have $GCD(a, b) = 1$. We need to show that  $a | c, b | c $ implies that  $ab | c$.

If $ GCD(a, b)= 1$ we have  some $m,n \in \Z$ such that $am + bn = 1$. If  $a | c, b | c$ then we have  some  $x,y \in \Z$ such that  $ax = c$ and  $by = c$.

We have, 
\begin{align*}
    am + bn &= 1\\
    cam + cbn &= c\\
    byam + axbn &= c \quad \text{ as $c = ax = by$}\\
    ab(ym + xn) &= c \\
    ab(z) &= c \\
\end{align*}
which by definition mean that $ab | c$


(b). Consider if $a = 5$ and  $b = 10$, so we have  $(a, b) = 5$.  We see that if $c = 20$ we have, $a = 5 | 20$ and  $b = 10 | 20$. However we see that $ab = 50 \not | 20$.


(c). We have $a_1, \dots, a_n \in \Z$ which are pairwise relatively prime numbers. We need to show that if $a_j | c$ then  $a_1\dots a_n | c$.

First we prove a primliminary result that given relatively prime numbers $a_1,\dots,a_n$, the gcd of the product of a subset of these numbers is relatively prime with numbers outside the subset. In other words we show that, 
$$ gcd(a_1\dots a_i, a_k) = 1 \text{ if $k > i$ } $$ 


We will do this by induction. For the base case we have $i = 1$  for which this is trivially true by construction (as all the numbers are pairwise relatively prime). Now consider the case for some arbitrary $m$. So we have,  
$$ gcd(a_1 \dots a_m, a_{m + 1}) = 1 $$ 

We need to show this is true for $m + 1$. Now let  $a_1 \dots a_m = x, a_{m + 1} = y, a_{k} = z$ where $k > m + 1$. Note that $x$ and  $y$ are coprime as well as $x$ and $z$ (by the inductive hypothesis).

So we have,

\begin{align*}
    xm_1 + zn_1 &= 1 \text{ for some $m_1,n_1 \in \Z$}\\
    ym_2 + zn_2 &= 1\text{ for some $m_2,n_2 \in \Z$}\\
\end{align*}

Now multiplying these two together we have, 
$$ xym_1m_2 + z(xm_1n_2 + zn_1n_2 + yn_1m_2) = 1 $$ 
$$ xym_3 + z(n_3) = 1 \text{ where $m_3 = m_1m_2, n_3 = xm_1n_2 + zn_1n_2 + yn_1m_2$} $$ 

This by definition means that $gcd(xy, z) = 1$. Or expending it gives us,

$$ gcd(a_1\dots a_{m}a_{m + 1}, a_{m + 2}) = 1 $$ 


Which is the case for $n = m + 1$


Now we prove the initial statement inductively.  Consider  $i = 1$ for which the statement is trivially true. Now consider the statement is true for some arbitrary $i$  so we have  $a_1\dots a_i | c$ given $a_1 | c,\dots,a_i | c$. Now, consider $a_{i + 1}$. Let $a_1 \dots a_i = a'$ so we have  $gcd(a', a_{i + 1}) = 1$ based on the proof above. Similarly we have $a' | c$ and $a_{i + 1} | c$ by assumption. So based on the proof in (a) this means that $a' a_{i + 1}| c$ or that $a_1 \dots a_{i + 1} | c$ which is the case for $i + 1$. Hence we compute the inductive step and show that it must be true for any  $i$.



\newpage 
\question  Exercise Set 1.3, \#48. Let $a,b\in \mathbb{Z} $. 
Show that $(a,b)=1$ if and only if $(a+b ,ab) =1$.  

\textbf{Solution.} 

We have $a, b \in \Z$. We need to show that  $(a, b) = 1$ if and only if  $(a + b, ab) = 1$. 

First we show the if condition. So we have, 
$$ (a + b, ab) = 1 $$  which means that, 
$$ (a + b) m + abn = 1 \text{ for some $m,n \in \Z$} $$ 

Now we have,
\begin{align*}
    am + bm + abn &= 1\\
    a(m + bn) + bm  &= 1\\
    ax + by &= 1
\end{align*}

By definition this means that $(a, b) = 1$


Now we show the only if condition.


Let's assume the contrary that $(a + b, ab ) = x > 1$ so we have,  
$$ x | a + b \text{ and } x | ab $$

If $x | ab$ let $p$ be a prime dividing $x$ then it must mean either  $p | a$ or $p | b$ as  $p$ can't divide  $a$ and  $b$ as   $gcd(a, b) = 1$. So assume without loss of generality that $ p | a$. Now we know that $x | a + b$ which means $p | a + b$. But if $p | a + b$ and $p | a$, then that must mean  $p$ also divides their difference or that $p | (a + b) - a$ or  $ p | b$. However, then we get that  $p | a$ and  $p | b$  or that $a$ and $b$ are not coprime as $p \ne 1$ which is a contradiction as we know that $gcd(a, b) = 1$. Hence, our assumption  must be wrong and it must be true that $(a + b, ab) = 1$. 




\newpage 
\question  Excercise Set 1.4, \#54(c).  Use the Euclidean Algorithm for this exercise. 
Compute the gcd $(441,1155) $, and express it as an integral linear combination of the $441$ and $1155$. 

\textbf{Solution.} 

We need to compute $gcd(441, 1155) $ using the euclidean algorithm,

 
\begin{align*}
    1155 &= 441 \times 2 +  273\\
    441 &= 273 \times 1 + 168\\
    273 &= 168 \times 1 + 105\\
    168 &= 105 \times 1 + 63\\
    105 &= 63 \times 1 + 42\\
    63 &= 42 \times 1 + 21\\
    42 &= 21 \times2 + 0
\end{align}
This gives us the GCD as $21$

To find the linear combination we go backwards to get,
\begin{align*}
    1 \times 105 &= 63 + (63 - 21)= 2\times63 - 21\\
    2 \times 168 &= 2 \times 105 + (105 + 21) = 3 \times 105 + 21\\
    3 \times 273 &= 3 \times 168 + (2 \times 168 - 21) = 5 \times 168 - 21\\
    5 \times 441 &= 5 \times 273  + (3 \times 273 + 21) = 8 \times 273 + 21\\
    8 \times 1155 &= 16 \times 441 + (5 \times 441 - 21) = 21 \times 441 - 21\\
\end{align*}
This gives us, 
    $$21\times 441 - 8 \times 1155 = 21$$






\newpage 
\question  Excercise Set 1.4, \#56. 
Find two rational numbers with denominators $11$ and $13$, whose sum is $\frac 7{143}$. 

\textbf{Solution.} 

We need to find two rations with denominators $11$ and  $13$ whose sum is  $\frac{7}{143}$. Consider the rationals to be $\frac{p}{11}$ and $\frac{q}{13}$ so we have, 
\begin{align*}
    \frac{p}{11} + \frac{q}{13} &= \frac{7}{143} \\
    \frac{13p + 11q      }{143} &= \frac{7}{143}\\
    13p + 11q &= 7
\end{align*}

We know that $13$ and  $11$ have gcd of  $1$ so there exists $m, n$ such that  $13m + 11n = 1$. 

 \begin{align*}
    13 &= 11 \times 1 + 2\\
    11 &= 2 \times 5 + 1\\
    5 \times 13 &= 5 \times 11  + (11 - 1)\\
    6 \times 11 -5 \times 13 &= 1
\end{align*}


Multiplying this by 7 we get, 
$$ 11 \times 42 + 13 \times -35= 7 $$ 


Hence we get $p = -35$ and $q = 42$ to get,  
$$ \frac{-35}{11} + \frac{42}{13} = \frac{7}{143} $$ 



\end{questions}

\end{document} 


 
%%%% don't delete the last line!
\end{document}
