
\section{Calculations}

\begin{eg}
    Compute a complete residue system mod $5$, 
    \begin{itemize}
        \item Using only even numbers
        \item Using only prime numbers
        \item Using only numbers congruent to $1 \pmod 4$
    \end{itemize}

    Default is $\{0, 1, 2, 3, 4\}$  so even numbers are $\{0, 6, 2, 8, 4\}$. For prime numbers we have,
    \begin{align*}
        0, 5\\ 
        1, 6, 11\\
        2, 7\\
        3, 8, 13\\
        4, 9, 14, 19
    \end{align*}

    So we have $\{5, 11, 7, 13, 19\} $
\end{eg}

\vspace{1em}
\hline
\vspace{1em}


\begin{note}
    
    We know that addition and multiplication are closed under congruence . We can think of this in terms of equivalnece classes, 
    \begin{align*}
        [a] + [b] &= [a + c]\\
        [b] \cdot [d] &= [bd]
    \end{align*}
    This turns the set of equivalence classes into a ring. We can construct addition and multiplication tables,
\end{note}

\begin{prop}
    Let $a, b, c, m\in \Z, m> 0$ then $ca \equiv c b \pmod m$ if and only if $a \equiv b \pmod {\frac{m}{(m, c)}}$
\end{prop}
\begin{proof}
    $\implies$. Assume $ca \equiv c b \pmod m$ so we have, $m \mid ca - c b$ or  $m \mid c( a - b)$. Let $d = (m, c)$. By transitivity we have  $\frac{m}{d} \mid \frac{c}{d} (a - b)$ but $(\frac{m}{d}, \frac{c}{d}) = 1$ which implies that $\frac{m}{d} \mid (a - b)$ or $ a \equiv b \pmod {\frac{m}{d}}$ by definition. 

    \vspace{1em}
    
    $\impliedby$. Assume  $a \equiv b \pmod {\frac{m}{(m,c)}}$ and $d = (m, c)$. We have  $\frac{m}{d} \mid  a -b$ so $m \mid d(a - b)$ and so  $m \mid d (a - b ) \frac{c}{d}$ or $m \mid c(a - b)$ or  $ca \equiv c b \pmod m$
\end{proof}

\section{Linear Congruences in one variable}%

\begin{definition}
    Let $a, b \in \Z$. A congruence of the form  $ax \equiv b \pmod m$ is called a \emph{linear congruence} in the variable  $x$.
\end{definition}

\begin{eg}
    If $2x \equiv 3 \pmod 4$  has no solutions. But  $2x \equiv 4 \pmod 6$ has  $x = 2$ as the only solution. And $3x \equiv 9 \pmod 6$ has  $1,3,5$.
\end{eg}

\begin{theorem}
    Let $ax \equiv b \pmod m$ and $d = (a, m)$. If  $d \nmid b$ then there are no solutions in $\Z$. Else, the congruence has exactly  $d$ \emph{incongruent} solutions modulo $m$ in  $\Z$.
\end{theorem}
\begin{note}
    This means that for any solution there are $d$ equivalence classes.
\end{note}
\begin{proof}
    Note that $ax \equiv b \pmod m$ iff $m \mid ax - b$ iff $ax - b = my$ for some $y \in \Z$ iff $ax - my = b$. Thus  $ax \equiv b \pmod m$ is solvable  in  $x$ if $ax  - my = b$ is solvable in $x,y$. Let $x, y$ be a solution of  $ax - my = b$. Since,  $d \mid a$ and  $d \mid m$ so  $d \mid b$. Taking contrapositives, if  $d \nmid b$  then there is no solution.
    \vspace{1em}

    Assume now that $d \mid b$. We prove  the second part in four steps.
    \begin{enumerate}
        \item We'll show that $ax \equiv b \pmod m$ has a solution  $x_0$.
        \item We'll show that there are infinitely many solutions of a particular form.
        \item We'll show that any solution has a particular form involving $x_0$ (combining with $2$ will give us all possible solutions).
        \item We'll show there are exactly $d$ equivalence classes.
    \end{enumerate}

    First, since $d = (a, m)$, there exists  $r, s \in \Z$ such that  $ar + ms = d$. Now as $d \mid b$ we have  $b = \frac{b}{d} d = \frac{b}{d} (ra + sm) = (\frac{b}{d}r) a + (\frac{b}{d}s) m$ thus $b - a(\frac{b}{d}) r = (\frac{b}{d}s) m$  and we have $m \mid b - a(\frac{b}{d}r)$. Thus $a(\frac{b}{d}r) \equiv b \pmod m$  and we have $x_0 = \frac{b}{d}r$ is a solution.


    \vspace{1em}
    
    Now, let $x_0$ be any solution. Consider the number $x_0 + (\frac{m}{d}) n$ where $n \in \Z$. So,  
    \begin{align*}
        a(x_0 + \frac{m}{d} n) &\equiv ax_0 +  \frac{m}{d}n \pmod m\\
                               &\equiv b + \frac{a}{d} m n \pmod m\\
                               & \equiv b \pmod m
    \end{align*}

    \vspace{1em}

    Let $x_0$ be an arbitrary solution of $ax \equiv b \pmod m$. So we have  $ax_0 - my_0 = b$ for some  $y_0 \in \Z$. Let $x$ be any other solution. Then $ax - my = b$ for some  $y \in \Z$. Subtracting both we have, 
    \begin{align*}
        (ax_0 - my_0) - (ax - my) &= 0\\
        a(x_0 - x) - m(y_0 - y) &= 0\\
        a(x_0 - x) &= m(y_0 - y)\\
        \frac{a}{d}(x_0 - x) &= \frac{m}{d}(y_0 - y)\\
    \end{align*}

    If $y_0 - y = 0$ then $x_0 - x = 0$. Now as solution are different we can assume $y_0 \ne y$. Now, we see that $(\frac{m}{d}, \frac{a}{d}) = 1$, so $\frac{m}{d} \mid \frac{a}{d}(x_0 - x)$ we have $\frac{m}{d} \mid x_0 - x$ by Prop 1.10. And we have $x \equiv x_0 \pmod {\frac{m}{d}}$. Thus, all solutions to $ax \equiv b \pmod m$ are given by $x = x_0 + \frac{m}{d} n, n \in \Z$ and $x_0$ is any particular solution.

    \vspace{1em}
    
    Let $x_0 + \frac{m}{d}n, x_0 + \frac{m}{d} n_2$ be solutions. Then, 
    \begin{align*}
        x_0 + \frac{m}{d} n_1 &\equiv x_0 + \frac{m}{d}n_2 \pmod m\\
         \frac{m}{d} n_1 &\equiv  \frac{m}{d}n_2 \pmod m\\
    \end{align*}

    This means that $m \mid \frac{m}{d}(n_1 - n_2)$ or $\frac{m}{d}(n_1 - n_2) = km$ and we have $n_1 - n_2 = kd$ and $n_1 \equiv n_2 \pmod d$. Since there are $d$ choices for the equivalence class of $n$. All solutions must fall into one of these cases.

\end{proof}










